---
title: "Analysing the COMPADRE database using R"
author: "Mario Vallejo-Marin"
date: '`r format(Sys.time(), "%B %d %Y")`'
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
    word_document: default
  classoption: a4paper
  word_document: default
---

#The COMPADRE database

Is a compilation of plant population matrices. See [www.compadre-db.org](http://www.compadre-db.org). Currently it has 7,024 population matrices for thousands of species.

##Downloading and reading the data in `R`

You can download the data from the website. The database is in `R` format. Save it to your computer, within the working environment of R, so you can then load it and analyse it.

```{r read.data, echo=T}
load(file="COMPADRE_v.4.0.1.RData")
```
This will result in a database object called `compadre`. It is a big file of about 45Mb.

The dataframe is arranged in a series of nested levels. At the highest level it has the following components:

```{r structure, echo=T}
summary(compadre)
```
The `metadata` section contains 47 variables. For a detailed description of the meaning of each of them see the user manual, which is available [here](https://github.com/jonesor/compadreDB/tree/master/COMPADRE-UserGuide).

Accessing the elements of `compadre` can be done using the operator `$`, just as you would do to access the columns of a simple data base.

```{r metadata, echo=T}
names(compadre$metadata)
```

With a bit of tweaking you can interrogate the database to obtain specific information. For example, to find ou how many genera are included in the data base you can do:

```{r numberofgenera, echo=T}
length(levels(as.factor(compadre$metadata$Genus)))
```

And you can print the first ten genera in this list:

```{r namesofgenera, echo=T}
levels(as.factor(compadre$metadata$Genus))[1:10]
```

##Data analysis

You can analyse subsets of the data using this general approach:

```{r subsets, echo=T}
#Subset the dataframe by Genus
mimulus.data<-subset(compadre$metadata, Genus=="Mimulus")
#Row numbers for the data containing Mimulus
keep<-as.numeric(rownames(mimulus.data))
#Subsets the matrices using the "keep" list
mat.mimulus<-compadre$mat[keep]
#This is a list of 36 matrices, which can be accessed using indices
mat.mimulus[[1]]
#And you can subset them further:
mat.mimulus[[1]]$matC
sum(mat.mimulus[[1]]$matC)
```

Now things more exciting, as we are ready to conduct analyses at a massive scale.

```{r analyses-loops, echp=T}
clonal.flag<-numeric()

#Loop to identify which matC entries are >0; i.e., have some measure of clonality
# Excludes indivisible matrices "NA"
n<-length(compadre$metadata$SpeciesAuthor)

for(i in 1:n){
  sum.matC<-sum(compadre$mat[[i]]$matC)
 # print(clonal.flag)
  if(!is.na(sum.matC) & sum.matC > 0) {
    clonal.flag<-c(clonal.flag,i)
    #print(cbind(i, clonal.flag))
    }
}

#List of species with matC > 0
compadre$metadata$SpeciesAccepted[clonal.flag][1:10]
length(compadre$metadata$SpeciesAccepted[clonal.flag])
levels(as.factor(compadre$metadata$SpeciesAccepted[clonal.flag]))[1:10]
```

For example, to calculate lambda:

```{r clonallambda, echo=T}
#Calculate lambda for all clonal species:
require(popbio)


n<-length(compadre$metadata$SpeciesAccepted[clonal.flag])
lambda.A<-rep(NA,n)

for(i in 1:n){
  tryCatch({  #Avoids error message due to missing values in x
  lambda.A[i]<-lambda(compadre$mat[[i]]$matA)
  }, error=function(e){})
}
sort(round(lambda.A,3))
hist(lambda.A, breaks="fd", xlim=c(0,7), col="black") #Omits from the plot all values of lambda>7

```