---
title: "Re-analysis Angert (2016)"
author: "Mario Vallejo-Marin"
date: "14/03/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Background
This exercise consists of a reanalysis of the datasets from Angert (2006).
The data is provided as part of the `popbio` package.

First load the library:

```{r load popbio, echo=TRUE}
library(popbio)
```

The dataset `monkeyflower` contains the transition rates of 32 populations of two *Mimulus* species, measured over multiple years. For an explanation of the transition rates see Angert (2006).

Here are the first few lines of the data set:

```{r monkeyflowerdata, echo=TRUE}
#Because Rmarkdown knits from a different environment, you need so save the data set first and then read it here. This is not necessary if you are working from the console.
#save(monkeyflower,file="monkeyflower.RData")
load(file="monkeyflower.RData")
head(monkeyflower)
```
The first three columns are species, site, and year. The data starting in the fourth column contains the transition rates.

To read the data in matrix form you need to get columns 4 to 19 for a given row. For instance, the data for the first row (*M. cardinalis*, population Buckmeadows, year 2000):

```{r mat1, echo=TRUE}
m1<-matrix(as.numeric(monkeyflower[1,4:19]), byrow=T, nrow=4)
m1
```

This produces a 4x4 matrix with 16 transition rates.

And from this you can easily conduct the population matrix analyses. For example to get `lambda`:

```{r lam1, echo=TRUE}
lambda(m1)
```
##Analysis over multiple matrices

Now you can do loops to analyse the properties of the 31 matrices at once.
```{r loops, echo=T}
temp.res<-data.frame(lambda=rep(NA,32), gentime=rep(NA,32))
for(i in 1:32){
  mattemp<-matrix(as.numeric(monkeyflower[i,4:19]), byrow=T, nrow=4)
  temp.res[i,1]<-lambda(mattemp) 
  temp.res[i,2]<- generation.time(mattemp)
}
par(mfrow=c(1,2))
hist(temp.res$gentime, breaks=10, main="Generation time", xlab="Generation time")
hist(temp.res$lambda, breaks=10, main="lambda", xlab="lambda")
```
